Make any stupid plot you want

### 2021-02-24 draft/sections/math.tex
math issue

The artist is the function that converts data into graphics; its name is taken from the analogous part of Matplotlib\cite{hunterArchitectureOpenSource} that builds visual elements to pass off to the renderer. The topological artist \vartist\  is a monoid equivariant sheaf map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation}

Usually \vartist can usually map into graphical elements solely with the data in \dsection, some visualizations, such as line, may also need some finite number $n$ of derivatives, which is captured by the jet bundle $\mathcal{J}^n$ \cite{JetBundle2020,musilovaCalculusVariationsJet2016} with $\mathcal{J}^{0}(\dtotal)=\dtotal$. In this work, we at most need $\mathcal{J}^{2}(\dtotal)$ which is the value at \dsection\ and its first and second derivatives; therefore the artist takes as input the data bundle padded with the jet bundle $\dtotal^{\prime} = \dtotal + \mathcal{J}^{2}(\dtotal)$. Specifically, \vartist is the map from $\dtotal^{\prime}$ to a specific graphic $\gsection \in \Gamma(H)$

\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
where the input can be point wise $\dsecton(\dbasepoint)\mid \dbasepoint \in \dbase$. The encoders $\vchannel:\dtotal^{\prime} \rightarrow \vtotal$ convert the data components to visual components(\ref{sec:graphic_base}). The continuity map $\vindex:\gbase \rightarrow \dbase$ then pulls back the visual bundle \vtotal\ over \gbase (\ref{sec:artist_nu}). Then the assembly function $\vmark: \vtotalpull \rightarrow \gtotal$ composites the fiber components of \vtotalpull\ into a graphic in \gtotal (\ref{sec:artist_q}). This functional decomposition of the visualization artists allows us to specify what are the responsibilities of each function in a fully constrained way. In turn, this allows for building reusable components at each stage of the transformation.

@@

 A section \vsection is a tuple of visual values that specifies the visual characteristics of a part of the graphic. For example, given a fiber of $\{xpos, ypos, color\}$ one possible section could be  $\{.5, .5, (255, 20,147)\}$. The $codomain(\vchannel_i)$ determines the monoid actions on $\vsection_i$. These fiber components are implicit in the library, by making them explicit as components of the fiber we can build consistent definitions and expectations of how these parameters behave.

\subsubsection{Visual Encoders \vchannel}
\label{sec:artist_nu}
%% equivariant maps 
\begin{figure}
    \includegraphics[width=\textwidth]{figures/math/equivariance_nu.png}
    \caption{In this artis, \vchannel\ maps the strings to the emojis. For \vchannel\ to be equivariant, a shuffle in the words should have an equivalent shuffle in the emojies, and a shuffle in the emojies should have an equivalent shuffle in the words.}
    \label{fig:artist_nu}
\end{figure}
As introduced in section~\ref{sec:intro_visual_variables}, there are many ways to encode data visually. We define the visual transformers \vchannel\ as the set of independent conversion functions 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
where $\vchannel_i: \dsection_i \mapsto \vsection_i$ is an equivariant map such that there is a monoid homomorphism from $\dfiber_{i}$ to $vfiber_{i}$. As mentioned in section~\ref{sec:data_monoid}, we choose monoid actions as the basis for equivariance because they define the structure on the fiber components. A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
    \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
commutes such that $\vchannel_i(m_{\delement}(\dtotal_i)) = m_{\velement}(\vchannel_i(\dtotal_i))$. This equivariance constraint yields guidance on what makes for an invalid transform. For example, the conversion $\vchannel_{i}(x) = .5$ does not commute under translation monoid action $t(x) = x+2$  
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + \vchannel(2)\\
    .5 &\neq .5 + .5
\end{align}

@@

As shown in figure~\ref{fig:artist_q}, the assembly function \vmark\ combines the fiber $\dfiber_i$ wise \vchannel\ transforms into a graphic in \gtotal. Together, \vchannel\ and \vmark\ are a map-reduce operation: map the data into their visual encodings, reduce the encodings into a graphic. As with \vchannel\, the constraint on \vmark\ is that for every monoid action on the input \vsection\, there is a corresponding monoid action on the output \gsection.

@@

The glyph in figure~\ref{fig:artist_mark_change} has the following characterstics \vfiber specified by  $(xpos,\, ypos,\, color,\, thickness)$ such that one section is $\vsection=(0,0,0,1)$ and $\vmark(\vsection) = \gsection$ generates a piece of the thin hollow circle. The equivariance constraint on \vmark is that the action $m=(e, e, e, x+2)$, where e is identity, applied to \vsection\ such that $\vsection^{\prime}=(e,e,e,3)$ has an equivalent action on \gsection that causes $\vmark(\vsection^{\prime})$ to be equivalent to the thicker circle in figure~\ref{fig:artist_mark_change}.

@@

\subsubsection{Sheaf and Stalk}
\label{sec:data_sheaf_stalk}
\note{move to artost}
Another use for the sheaf is to provide access to all the records over the space on which the sheaf $\mathcal{O}(\dtotal)$ is defined. The artist \vartist\ could not generate any visualization that needed more information than was in a single record, such as a line plot, if it took as input \dsection, because $\dsection:\dbasepoint \mapsto \delement$ only knows about the record it returns. 
The data bundle \dtotal\ is defined globally on \dbase, but often we want as

 To make these types of visualizations, mathematical objects called sheafs provide a way to construct a \dsection\ on a subset of \dbase.

As part of the definition of local triviality, there is an open neighborhood $U\subset \dbase$ for every $\dbasepoint \in \dbase$. We can define the inclusion map $\iota: U \rightarrow \dbase$  which pulls \dtotal over $U$ 
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$. Restriction maps on the sheaf define how to glue together these localized sections $\iota^*\dsection$ into larger sections\cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018} over the subset of \dbase\ requested by the visualization.

The sheaf facilitates visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995} streaming data, or navigation such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006}.; the jet ensures that the artist only needs a single record on \dbasepoint\ extended with derivatives to render a piece of a graphic. This allows for the creation of highly concurrent and possibly distributed artists.

@@

\subsubsection{Equivalance class of artists $\vartist^{\prime}$}
\label{sec:artist_equivalance}
It is impractical to implement an artist for every single graphic; instead we implement the equivalence class of artists $\{\vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}\}$ which is essentially an artist where the \vchannel\ functions have not yet been input. Equivalent artists have the same fiber bundle \vtotal\ and same assembly function \vmark\, but act on different sections \vsection. To further simplify implementation, we identify a minimal \vfiber\ associated with each $\vartist^{\prime}$ that defines what visual characteristics of the graphic must originate in the data such that the graphic is identifiable as a given chart type.

>>>>

The artist is the function that converts data into graphics; its name is taken from the analogous part of Matplotlib\cite{hunterArchitectureOpenSource} that builds visual elements to pass off to the renderer. The topological artist \vartist\  is a monoid equivariant sheaf map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation}
Sheafs are a mathematical object with restriction maps that define how to glue \dsection over local neighborhoods $U \subseteq \dbase$,  discussed in section~\ref{sec:artist_sec:data_sheaf_stalk}, such that the \vartist\ maps are consistent over continuous regions of \dbase. While \vartist can usually construct graphical elements solely with the data in \dsection, some visualizations, such as line, may also need some finite number $n$ of derivatives, which is captured by the jet bundle $\mathcal{J}^n$ \cite{JetBundle2020,musilovaCalculusVariationsJet2016} with $\mathcal{J}^{0}(\dtotal)=\dtotal$. In this work, we at most need $\mathcal{J}^{2}(\dtotal)$ which is the value at \dsection\ and its first and second derivatives; therefore the artist takes as input the data bundle padded with the jet bundle $\dtotal^{\prime} = \dtotal + \mathcal{J}^{2}(\dtotal)$. Specifically, \vartist is the map from $\dtotal^{\prime}$ to a specific graphic $\gsection \in \Gamma(H)$

\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
where the input can be point wise $\dsection(\dbasepoint)\mid \dbasepoint \in \dbase$. The encoders $\vchannel:\dtotal^{\prime} \rightarrow \vtotal$ convert the data components to visual components(\ref{sec:graphic_base}). The continuity map $\vindex:\gbase \rightarrow \dbase$ then pulls back the visual bundle \vtotal\ over \gbase (\ref{sec:artist_nu}). Then the assembly function $\vmark: \vtotalpull \rightarrow \gtotal$ composites the fiber components of \vtotalpull\ into a graphic in \gtotal (\ref{sec:artist_q}). This functional decomposition of the visualization artists allows us to specify what are the responsibilities of each function in a fully constrained way. In turn, this allows for building reusable components at each stage of the transformation.

@@

 A section \vsection is a tuple of visual values that specifies the visual characteristics of a part of the graphic. For example, given a fiber of $\{xpos, ypos, color\}$ one possible section could be  $\{.5, .5, (255, 20,147)\}$. The $codomain(\vchannel_i)$ determines the monoid actions on $\vfiber_i$. These fiber components are implicit in the library, by making them explicit as components of the fiber we can build consistent definitions and expectations of how these parameters behave.

\subsubsection{Visual Encoders \vchannel}
\label{sec:artist_nu}
%% equivariant maps 
\begin{figure}
    \includegraphics[width=\textwidth]{figures/math/equivariance_nu.png}
    \caption{In this artis, \vchannel\ maps the strings to the emojis. For \vchannel\ to be equivariant, a shuffle in the words should have an equivalent shuffle in the emojies, and a shuffle in the emojies should have an equivalent shuffle in the words.}
    \label{fig:artist_nu}
\end{figure}
As introduced in section~\ref{sec:intro_visual_variables}, there are many ways to encode data visually. We define the visual transformers \vchannel\ 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in section~\ref{sec:data_monoid}, we choose monoid actions as the basis for equivariance because they define the structure on the fiber components.

A validly constructed \vchannel\ is one where the diagram of the monoid transform $m$
\begin{equation}
    \label{eq:nu_categorical}
\begin{tikzcd}
    \dtotal_i \arrow[r] \arrow[r, "\vchannel_i"] \arrow[d, "m_{\delement}"'] & \vtotal_i \arrow[d, "m_{\velement}"] \\
    \dtotal_i \arrow[r, "\vchannel_i"]                           & \vtotal_i               
\end{tikzcd}
\end{equation}
commutes such that $\vchannel_i(m_{\delement}(\dtotal_i)) = m_{\velement}(\vchannel_i(\dtotal_i))$. This equivariance constraint yields guidance on what makes for an invalid transform. For example, the conversion $\vchannel_{i}(x) = .5$ does not commute under translation monoid action $t(x) = x+2$  
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + \vchannel(2)\\
    .5 &\neq .5 + .5
\end{align}

@@

As shown in figure~\ref{fig:artist_q}, the assembly function \vmark\ combines the fiber $\dfiber_i$ wise \vchannel\ transforms into a graphic in \gtotal. Together, \vchannel\ and \vmark\ are a map-reduce operation: map the data into their visual encodings, reduce the encodings into a graphic. As with \vchannel\, the constraint on \vmark\ is that for every monoid action on the input \vsection\, there is corresponding monoid action on the output \gsection.

@@

The glyph in figure~\ref{fig:artist_mark_change} has the following characteristics \vfiber specified by  $(xpos,\, ypos,\, color,\, thickness)$ such that one section is $\vsection=(0,0,0,1)$ and $\vmark(\vsection) = \gsection$ generates a piece of the thin hollow circle. The equivariance constraint on \vmark\ is that the action $m=(e, e, e, x+2)$, where e is identity, translates \vsection\ to  $\vsection^{\prime}=(e,e,e,3)$. The corresponding action on \gsection causes $\vmark(\vsection^{\prime})$ to be the thicker circle in figure~\ref{fig:artist_mark_change}.

@@

\subsubsection{Sheaf}
\label{sec:artist_sheaf_stalk}
As part of the definition of local triviality, there is an open neighborhood $U\subset \dbase$ for every $\dbasepoint \in \dbase$. We can define the inclusion map $\iota: U \rightarrow \dbase$  which pulls \dtotal over $U$ 
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
such that the pulled back $\iota^*\dsection$ only contains records over $U \subset \dbase$. Restriction maps on the sheaf define how to glue together these localized sections $\iota^*\dsection$ into larger sections\cite{ghristElementaryAppliedTopology2014,ghristHomologicalAlgebraData2018} over the subset of \dbase\ requested by the visualization.

The sheaf facilitates visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995} streaming data, or navigation such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006}.;

@@

\subsubsection{Equivalance class of artists $\vartist^{\prime}$}
\label{sec:artist_equivalance}
It is impractical to implement an artist for every single graphic; instead we implement an approximation of an the equivalence class of artists $\{\vartist \in \vartist^{\prime}: \vartist_{1} \equiv \vartist_{2}\}$. Roughly, equivalent artists have the same fiber bundle \vtotal\ and same assembly function \vmark\ but act on different sections \vsection, but we will formalize the definition of the equivalance class in future work. As a first pass for implementation purposes, we identify a minimal \vfiber\ associated with each $\vartist^{\prime}$ that defines what visual characteristics of the graphic must originate in the data such that the graphic is identifiable as a given chart type.

### 2021-02-24 draft/sections/topodata.tex
math issue

%%pull out cause not helping
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/triangle_torus.png}
    \caption{The torus \dtotal is unraveled into a simplacial complex of 2 faces \dbase. Transition functions are defined on the edges of \dbase such that surface can be glued back into the torus.
    \note{add cross sections a and b to ring and color same as edges in complex}}
    \label{fig:triangle_torus}
\end{figure}
An advantage of fiber bundles is the ability to encode data continuity that is more complex than points, a line or a plane. For example, the tori in figure~\ref{fig:triangle_torus} can be unraveled into a simplacial complex of two triangles with labeled edges. Transition functions $\delta$ are defined on the edges such that $a$ can be glued to $a^\prime$ and $b$ to $b^\prime$ to reconstruct the torus. This simplacial complex is then used as the base space encoding the continuity of data that lies in the torus. A constraint on the transition functions is that the monoid actions on the fibers on the edges of \dtotal\ are commutative $\monoid*\dfiber\mapsto \delta(\monoid\dfiber)= \monoid*\delta(\dfiber)$

>>>>

\subsection{Making the fiber bundle computable}
\label{sec:triangulization}
\begin{figure}[H]
    \includegraphics{figures/math/simplex.png}
    \caption{Simplices can encode the connectivity of the data, from fully disconnected (0 simplex) records to all records are connected to at least 3 others}
    \label{fig:triangle_simplex}
\end{figure}
%%big data
One way of expressing the connectivity of records in a dataset is to implement \dbase\ as a simplacial complex, which is a set of simplicies such as those shown in figure~\ref{fig:triangle_simplex}. The advantage of triangulization is that it is general enough to work for more complex topology based visualization methods \cite{heineSurveyTopologybasedMethods2016} while also providing a consistent interface of vertices, edges, and faces for \vindex\ to map into. When triangulated, the simplices encode the continuity in the data

\begin{table}[H]
    \begin{center}
        \begin{tabular}{|l|l|l|}\hline
        \textbf{simplex} & \textbf{continuity} & \textbf{\dsection}   \\ \hline
        vertex  & discrete   & \dsection(\dbasepoint)                  \\ \hline
        edge    & 1D         & \dsection(\dbasepoint,\, \dx)        \\ \hline
        face    & 2D         & \dsection(\dbasepoint,\, \dx,\, \dy)\\ \hline
        \end{tabular}
        \caption{}
        \label{tab:triangulization}
    \end{center}
\end{table}

such that each section is bound to a simplex $\dbasepoint \in \dbase$. As shown in table~\ref{tab:triangulization}, in a 1D continuous spaces each \dsection\ lies distance \dx\ along edge \dbasepoint, while in a 2D continuous space each \dsection\ lies at coordinate \dx,\dy on the face \dbasepoint. This is directly analogous to indexing to express connectivity in N-D arrays, while also natively supporting graphs and trees as they are simplacial complices of nodes and edges. Path connected components are then sections where edges or faces meet.

In general simplacial complexes, in these prototypes we implement graphs. 
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/triangle_torus.png}
    \caption{The torus \dtotal is unraveled into a simplacial complex of 2 faces \dbase. Transition functions are defined on the edges of \dbase such that surface can be glued back into the torus.
    \note{add cross sections a and b to ring and color same as edges in complex}}
    \label{fig:triangle_torus}
\end{figure}
An advantage of fiber bundles is the ability to encode data continuity that is more complex than points, a line or a plane. For example, the tori in figure~\ref{fig:triangle_torus} can be unraveled into a simplacial complex of two triangles with labeled edges. Transition functions $\delta$ are defined on the edges such that $a$ can be glued to $a^\prime$ and $b$ to $b^\prime$ to reconstruct the torus. This simplacial complex is then used as the base space encoding the continuity of data that lies in the torus. A constraint on the transition functions is that the monoid actions on the fibers on the edges of \dtotal\ are commutative $\monoid*\dfiber\mapsto \delta(\monoid\dfiber)= \monoid*\delta(\dfiber)$
